---
title: "Lab_13"
author: "Lil Rascals"
date: "April 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(cowplot)
library(stringr)
library(gapminder)
```
### Overall Question: What does world health look like?
This is interesting, because it is essential to note what health is like in your country and countries around the world. We could use this to see what measures should be taken in different countries to improve health. 

#### Specific Question Kelsey
Is health or more specifically smoking in women better in the US than other countries? 
```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(stringr)
library(gapminder)
library(forcats)
library(cowplot)
smoke <- read_csv("smoking_women.csv")



smoke<-na.omit(smoke) %>% gather('2000', '2002','2005', key = "year", value = "percent")
smokepercent <- function(perms = 1000, year1,country1, n1 ){


  mean_mat <- matrix(0,1000,1)
  for (i in c(1:perms))
  {

    cont1 <- smoke %>% filter(country==country1, year==year1) 
    cont2 <- smoke %>% filter(year==year1) 
    other <- sample(cont2$percent, n1, replace = FALSE)

    A <- mean(cont1$percent)
    B <- mean(other)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}

rsample <- as.tibble(smokepercent(perms = 1000,year1=2005,country1="United States",n1=10))

cont3 <- smoke %>% filter(year==2005, country=="United States") %>% select(percent)
cont4 <- smoke %>% filter(year==2005) %>% select(percent)
smoke1 <- summary(cont3$percent)
smoke2 <- summary(cont4$percent)
diff <- median(cont3$percent)- median(cont4$percent)

std <- sqrt( (sd(cont4$percent))^2/length(cont4$percent))
z_score <- diff/std
z_score

ggplot(rsample, aes(x=V1 ))+geom_histogram()+
  xlab("Means")+
  geom_vline(xintercept = diff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Women Smoking in 2005")

```
####Results Kelsey McKenna
I found that there is clear evidence that more people in the United States smoked in 2005 compared with other countries. This data is statistically significant at nearly every level. This shows that we have 11.75% more women who smoke in the United States which may indicate that we could be less healthy and more prone to lung disease than other countries around the world. 

####Specific Question Shreeya

```{r}
aids <- read_csv("aids_deaths.csv")
aids<-na.omit(aids) %>% gather('2000', '2001','2005', key = "year", value = "number")%>% group_by(year)
##View(aids)
cont2 <- aids %>% filter(year=="2001")%>% select(number)

   other <- sample(cont2$number, 20, replace = FALSE)


aidsnum <- function(perms = 1000, year1,country1, n1 ){


  mean_mat <- matrix(0,1000,1)
  for (i in c(1:perms))
  {

    cont1 <- aids %>% filter(country==country1, year==year1) 
    cont2 <- aids %>% filter(year==year1) 
    other <- sample(cont2$number, n1, replace = FALSE)

    A <- mean(cont1$number)
    B <- mean(other)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}

rsample <- as.tibble(aidsnum(perms = 1000,year1=2000,country1="United States", n1=10))
cont3 <- aids %>% filter(year==2001, country=="United States") %>% select(number)
cont4 <- aids %>% filter(year==2005) %>% select(number)

aids1 <- summary(cont3$number)
aids2 <- summary(cont4$number)
diff <- median(cont3$number)- median(cont4$number)

std <- sqrt( (sd(cont4$number))^2/length(cont4$number))
z_score <- diff/std
z_score

ggplot(rsample, aes(x=V1 ))+geom_histogram()+
  xlab("Means")+
  geom_vline(xintercept = diff, color = "orange", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means of Deaths due to AIDS in 2001")

```
Findings: I found that statistically, the United States had a higher mean of deaths from AIDS in 2001. This is incredibly important information because AIDS is manageable with the proper medicine and so higher rates of AIDs related deaths is an indicator of the health of a country and it's accessibility to medicine. This information is especially crucial at a time where HIV research is rapidly advancing. As is shown in the histogram, the United States has a higher mean than almost all countries, although the United States has comparable deaths to many countries in Africa, where HIV is also a big health issue.

####Specific Question Matthew
Does health spending have an effect on the overall health of a country?


```{r}
#Health spending per person
HSPP <- read_csv("percap.csv", na = "NA")

#arrange by year which country spends the least and most
HSPP1 <- HSPP %>% gather(key = Year, SPC, `1995.0`:`2010.0`) %>% group_by(Year)
HSPP <- HSPP1 %>% transform(Year = as.numeric(HSPP1$Year), SPC = as.numeric(HSPP1$SPC)) %>% na.omit()
HSPP %>% arrange(desc(SPC))
names(HSPP) <- c("Country", "Year", "SPC")


Spending_per_capita <- function(perms = 1000, country1 ,  n1 ){
  mean_mat <- matrix(0,1000,1)
  for (i in c(1:perms))
  {
    cont5 <- HSPP %>% filter(Country == country1) 
    cont6 <- HSPP 

    rcont5 <- sample(cont5$SPC, n1, replace = FALSE)
    rcont6 <- sample(cont6$SPC, length(cont6$SPC)-n1, replace = FALSE)
    
    A <- mean(rcont5)
    B <- mean(rcont6)
    
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
rsample <- as.tibble(Spending_per_capita(perms = 1000, country1 = "United States",  n1 = 15))
this <- HSPP %>% filter( Country=="United States") %>% select(SPC)
this1 <- HSPP %>% select(SPC)
diff <- mean(this$SPC)- mean(this1$SPC)
ggplot(rsample, aes(x=V1 ))+geom_histogram()+
  xlab("Means")+
  geom_vline(xintercept = diff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Years of Spedings Means")
```

#### Specific Question James
Is health or more specifically smoking in women better in the US than other countries? 
```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(stringr)
library(gapminder)
library(forcats)
library(cowplot)
bp <- read_csv("blood_pressure.csv")

bp1 <- bp %>% gather(key = Year, SBP, `1980`:`2008`) 
bp2 <- bp1 %>% transform(Year = as.numeric(bp1$Year), SPC = as.numeric(bp1$SBP)) %>% na.omit()
bp2 %>% arrange(desc(SBP))
names(bp2) <- c("Country", "Year", "SBP")

Blood_pressure <- function(perms = 1000, country1  ){
  mean_mat <- matrix(0,1000,1)
  for (i in c(1:perms))
  {
    cont8 <- bp1 %>% filter(country == country1) %>% select(SBP)
    cont9 <- bp1 %>% select(SBP)

    rcont9 <- sample(cont9$SBP, 15, replace = FALSE)
    
    A <- mean(cont8$SBP)
    B <- mean(rcont9)
    
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
s <- as.tibble(Blood_pressure(perms = 1000,country1 = "United States"))

ggplot(s, aes(x=V1 ))+
  geom_histogram()+
  xlab("Means")+
  geom_vline(xintercept = diff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Men's Blood Pressure")
```
  
####Results James Burke


###Summary
####Kelsey McKenna
I observed the differences in women who smoke between the US and other countries and conducted a difference of means permutation test to test this claim. I found that women are much more likely to smoke than in other countries. I used the test we learned in class and a function so you could change the year easily and see how it changed or change which country you are comparing very easily. I had to gather the data to get the years in a format that was easier to work with. It had countries years and percent who smoke. 

